---
title: "Financial Time Series Models"
---   
Generalized AutoRegressive Conditional Heteroskedasticity (GARCH) is a statistical model used in analyzing time-series data where the variance error is believed to be serially autocorrelated. The term heteroskedasticity describes the irregular pattern of variation of an error term, or variable, in statistical model. GARCH models assusme that the variance of the error term follows an autoregressive moving average process. GARCH is useful to assess risk and expected returns for assests that exhibit clustered periods of volatility in returns.  

In the following models, we are going to analyze the ExxonMobil stock returns by using ARMA+GARCH models and then predict the future returns volitility.

```{r, echo=FALSE,message=FALSE,warning=FALSE}
library(tidyverse)
library(ggplot2)
library(forecast)
library(astsa) 
library(xts)
library(tseries)
library(fpp2)
library(fma)
library(lubridate)
library(tidyverse)
library(TSstudio)
library(quantmod)
library(tidyquant)
library(plotly)
library(fGarch)
```
# ARMA+GARCH Models for ExxonMobil Stock Returns
## Explore the Data 

Let's first explore the data of the returns of ExxonMobil (XOM) stock prices from January 1985 to April 2023. The original data, XOM stock prices, is extracted from Yahoo Finance, where the plot below is the returns of the adjusted closing stock prices for XOM, which is calculated by taking the log of the XOM stock prices and take the first difference. From the plot, we know that there are clear seasonality and great volatility of returns of the stock prices in these 38 years. This data is suitable for building GARCH model.  

```{r, echo=FALSE}
xom <- read.csv("./data/yahoo/energy/XOM.csv")
xom$Date <- as.Date(xom$Date, "%Y-%m-%d")
xom <- subset(xom, select=c(Adj.Close))
xom.ts <- ts(xom, start=c(1985,1), end=c(2023,4), frequency = 12)
xom.returns <- log(xom.ts) %>% diff()
#head(xom)
#tail(xom)
autoplot(xom.returns, col=2, xlab="Year", ylab="Returns", main="ExxonMobil Returns (Closing Adjusted Stock Price) from 1985 to 2023")
```
Let's first look at the ACF and PACF plots for the returns of ExxonMobil closing stock prices. Both look pretty good as almost all lags are within the confidence bands.  
```{r, echo=FALSE, message=FALSE}
require(gridExtra)
p1 <- ggAcf(xom.returns,40)
p2 <- ggPacf(xom.returns, 40)
grid.arrange(p1,p2, nrow=1)
```

However, when looking at the ACF and PACF plots for the absolute values and squared values of the returns, all appear to have correlations in seasonality property, which too many lags are out of the bands and clear seasonal components exist. Therefore, fitting an ARIMA model first would be more appropriate.  

```{r, echo=FALSE}
p1 <- ggAcf(abs(xom.returns), 40)
p2 <- ggAcf(xom.returns^2, 40)
p3 <- ggPacf(abs(xom.returns), 40)
p4 <- ggPacf(xom.returns^2, 40)
grid.arrange(p1,p2, p3, p4, nrow=2)
```

## Fitting ARIMA+GARCH Models
### Fitting ARIMA Model

```{r, echo=FALSE}
ARIMA.c=function(p1,p2,q1,q2,data){
temp=c()
d=1
i=1
temp= data.frame()
ls=matrix(rep(NA,6*50),nrow=50)


for (p in p1:p2)
{
  for(q in q1:q2)
  {
    for(d in 0:2)
    {
      
      if(p+d+q<=6)
      {
        
        model<- Arima(data,order=c(p,d,q))
        ls[i,]= c(p,d,q,model$aic,model$bic,model$aicc)
        i=i+1
  
        
      }
      
    }
  }
}


temp= as.data.frame(ls)
names(temp)= c("p","d","q","AIC","BIC","AICc")

temp
}
```

```{r, echo=FALSE}
output <- ARIMA.c(1,3,1,3,data=xom.returns)
#output
```

Now, we are fitting an ARIMA model by using appropriate p, d, and q. From the previous original ACF and PACF plots, it can be decided that p=1,3, and q=1,3, where these are significant values on both plots, and then we can fit an ARIMA model within this range for p and q for a series of models, and find the best one for the minimum AIC, AICc, and BIC values.  

From the output with minimum AIC, AICc and BIC values below, only one model is provided. Therefore, the ARIMA(1,0,1) or ARMA(1,1) would be the best model.  

```{r, echo=FALSE}
aic <- output[which.min(output$AIC),]
aicc <- output[which.min(output$AICc),]
bic <- output[which.min(output$BIC),]

mintable <- rbind(aic, aicc, bic)
row.names(mintable) <- c("Minimum AIC", "Minimum AICc", "Minimum BIC")
knitr::kable(mintable)
```

Below are the plots for the residuals of the ARMA model. When looking closer at the residuals of the model with drift, there are clear volatility and seasonality in the residuals, therefore, further modelling is needed on this returns data for ExxonMobil.  

```{r,echo=FALSE}
fit <- Arima(xom.returns, order=c(1,0,1),include.drift = TRUE)
#summary(fit)
fit.re <- fit$residuals
checkresiduals(fit)
```
### Fitting GARCH Model

Here are the ACF and PACF plots for the squared residuals of the above ARMA model. Both plots show a large number of lags are out of the significance bands, indicating a high auto-correlation structure. It is evident that GARCH(1,1) model would be a fit for the data, because they are too high to be inside of the bands. From the lags, we can decide that both p and q values are from 1 to 5. Next, let's fit the GARCH model.  

```{r, echo=FALSE}
p1 <- ggAcf(fit.re^2)
p2 <- ggPacf(fit.re^2)
grid.arrange(p1,p2, nrow=1)
```

From the result of fitting a GARCH model on the ARMA residuals below, the `GARCH(1,2)` is the best model, which having then lowest AIC score from the overall GARCH model evaluation. 
```{r, echo=FALSE}
model <- list() ## set counter
cc <- 1
for (p in 1:5) {
  for (q in 1:5) {
  
model[[cc]] <- garch(fit.re,order=c(q,p),trace=F)
cc <- cc + 1
}
} 

## get AIC values for model evaluation
GARCH_AIC <- sapply(model, AIC) ## model with lowest AIC is the best
#which(GARCH_AIC == min(GARCH_AIC))
model[[which(GARCH_AIC == min(GARCH_AIC))]]
```
Below are the summary statistics for both the ARMA model and the GARCH model on the residuals of the ARMA model, which can be used for writing model equations later.  

```{r, echo=FALSE}
summary(fit)
g.summary <- summary(fit12 <- garchFit(~garch(1,2),fit.re,trace = F))

```

### Model Diagnostic
Our best model for here right now is `ARMA(1,1)+ GARCH(1,2)`. Let's first check the residuals of the GARCH model. It has done a relatively good job, even though not a huge improvement from the ARMA(1,1) model diagnostic alone, while the ACF and the shape of the residuals are pretty good.  

```{r, warning=FALSE, echo=FALSE}
fit2 <- garch(fit.re,order=c(1,2),trace=F)
checkresiduals(fit2) #relatively doing a good job
```
Moreover, there isn't a significant deviation from the straight line, only a slightly deviation for both tails, upper and lower. The distribution is fairly even, as there are not showing more heavy-tailed of returns than normal distributions from the two tails.  

```{r, echo=FALSE}
qqnorm(fit2$residuals, pch = 1)
qqline(fit2$residuals, col = "blue", lwd = 2)
```
From the Box-Ljung test, the p-value is significantly greater than 0.05, while approximately to 1. The Ho hypothesis states that the data are independently distributed (the correlations in the population from which the sample is taken are 0, so that any observed correlations in the data result from randomness of the sampling process); while the Ha hypothesis states that the data are not independent distributed; they exhibit serial correlation. Since the p-value is so large, we do not have enough evidence to reject the Ho hypothesis, which is a desired result, because we look for to say no correlation left. Therefore, our two models have done a fairly good job on the returns of Exxon Mobil stock prices data.  

```{r, echo=FALSE}
Box.test(fit2$residuals, type = "Ljung")
```
### Equations for the Best Model

From the two summary statistics above, the equations for the best model `ARMA(1,1) + GARCH(1,2)` can be written as: 

$$x_t = 0.3166x_{t-1} +w_t-0.35473w_{t-1}+ 0.0143$$
$$z_t=\sigma_t \epsilon_t$$ 

$$\sigma_t = 0.00012 + 0.1043z_{t-1}+ 0.057 \sigma_{t-1} +0.803\sigma_{t-2}$$


### Forecasting
Below is the forecasting plot for the GARCH(1,2) model with a 95% confidence interval. We see that the predicted returns for Exxon Mobil is not going to increase or decrease a lot, and are pretty stable in the near future by our model.  

```{r, echo=FALSE}
predict(fit12 , n.ahead = 20, trace = FALSE, plot=TRUE)
```


# Works Cited

THE INVESTOPEDIA TEAM, June 11, 2021, What Is Generalized AutoRegressive Conditional Heteroskedasticity (GARCH)?, Investopedia, https://www.investopedia.com/terms/g/garch.asp#:~:text=Generalized%20AutoRegressive%20Conditional%20Heteroskedasticity%20(GARCH)%20is%20a%20statistical%20model%20used,an%20autoregressive%20moving%20average%20process.














